---
title: "Model Card: H2O Stacking Ensemble"
date: '`r Sys.Date()`'
output: 
  html_document
params:
    name: conformalised_h2o_stack
    version: NULL
---

```{r, echo = FALSE, output=TRUE, eval=TRUE}
library(vetiver)
library(pins)
library(yardstick)
knitr::opts_chunk$set(echo = FALSE)

board <- board_s3(
  "data", 
  access_key = "user", 
  secret_access_key = "password", 
  region = "us-east-2", 
  endpoint = "https://minio-api.k8s.dev.co.nz"
)

v <- vetiver_pin_read(board, params$name, version = params$version)
v_meta <- pin_meta(board, params$name)

box::use(
  dials[svm_margin, trees],
  ggplot2[theme_set, theme_light],
  readr[read_csv],
  tune[control_bayes, extract_parameter_set_dials],
  here[here],
  ranger[ranger],
  kernlab,
  rsample[initial_split, vfold_cv],
  parsnip[rand_forest, svm_linear],
  dplyr[mutate],
  recipes[recipe, update_role, step_zv, step_nzv, step_normalize, step_impute_knn, update],
  yardstick[rmse, mae],
  hardhat[tune],
  workflows[workflow],
  workflowsets[workflow_set, option_add, option_add_parameters, workflow_map],
  stacks[stacks, add_candidates, blend_predictions, fit_members],
  vetiver[vetiver_model, vetiver_pin_write, vetiver_api],
  pins[board_temp, board_s3, pin_write, pin_read, pin_upload, pin_download],
  plumber[pr, pr_run],
  h2o[h2o.connect],
  agua[extract_fit_parsnip, rank_results, member_weights]
)

h2o.connect(
  ip = "h2o.k8s.dev.co.nz", port = 443, https = TRUE
)

theme_set(theme_light())

```

A [model card](https://doi.org/10.1145/3287560.3287596) provides brief, transparent, responsible reporting for a trained machine learning model.

## Model details

-   Developed by PERSON AND/OR TEAM
-   `r cli::pluralize("{v$description} using {ncol(v$prototype)} feature{?s}")`
-   More details about how model was developed and what it is predicting
-   More details on feature engineering and/or data preprocessing for model
-   Version `r v$metadata$version` of this model was published at `r v_meta$created`
-   Citation and/or license details for the model
-   If you have questions about this model, please contact [PERSON\@ORG.ORG](mailto:PERSON@ORG.ORG){.email}

## Intended use

-   The primary intended uses of this model are ...
-   The primary intended users of this model are ...
-   Some use cases are out of scope for this model, such as ...

## Important aspects/factors

-   Aspects or factors (demographic, environmental, technical) that are relevant to the context of this model are ...
-   In evaluating this model, we examined aspects such as ...

## Metrics

-   The metrics used to evaluate this model are ...
-   These metrics are computed via ...
-   We chose these metrics because ...

## Training data & evaluation data

-   The training dataset for this model was ...

-   The training dataset for this model has the "prototype" or signature:

```{r, eval=TRUE}
dplyr::glimpse(v$prototype)
```

-   The evaluation dataset used in this model card is ...

-   We chose this evaluation data because ...

```{r, eval=TRUE}
## EVALUATION DATA:

df <- v$model$training

## consider using a package like skimr or DataExplorer for automated 
## presentation of evaluation data characteristics
```

## Quantitative analyses {.tabset}

```{r, eval=TRUE}
## compute predictions for your evaluation data
## load packages needed for prediction:
library(parsnip)
library(workflows)
library(probably)
preds <- predict(v, df) |>
  dplyr::bind_cols(df)

preds  
```

### Overall model performance

```{r, eval=TRUE}
preds |>
    metrics(mpg, .pred)
```

### Visualize model performance

```{r, fig.height=3, eval=TRUE}
library(ggplot2)
p <- preds |>
    ggplot(aes(mpg, .pred)) +
    geom_abline(slope = 1, lty = 2, color = "darkblue", size = 1.2) +
    geom_point(alpha = 0.5, show.legend = FALSE) +
    geom_ribbon(aes(ymin = .pred_lower, ymax = .pred_upper), fill = "lightblue", alpha = 0.2)

plotly::ggplotly(p)
```

## Ethical considerations

-   We considered ...

## Caveats & recommendations

-   This model does ...
-   This model does not ...
-   We recommend ...
